---
title: "Data Challenge 2026"
format:
  html:
    theme: cyborg
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# Librerias a utilizar
library(dplyr)
library(ggplot2)
library(plotly)

# Lectura de los data sets
procurement <- read.csv("Datos modificados/procurement_modificado.csv")
catalogo_especificaciones <- read.csv("Datos modificados/catalogo_especificaciones.csv")
catalogo_operaciones <- read.csv("Datos modificados/catalogo_operaciones.csv")
catalogo_especificaciones_operaciones <- read.csv("Datos modificados/catalogo_especificaciones_operaciones.csv")
catalogo_especificaciones_procurement <- read.csv("Datos modificados/catalogo_especificaciones_procurement.csv")
especificaciones_procurement <- read.csv("Datos modificados/especificaciones_procurement.csv")

```

```{r}
# Porcentajes de utilización del pallet
cuadrados_porcentaje <- function(porcentajes, etiquetas = NULL, espacio = 0.4, titulo = "") {
  
  n <- length(porcentajes)
  if (is.null(etiquetas)) etiquetas <- paste0("Item ", 1:n)
  
  shapes_list <- list()
  annotations_list <- list()
  
  for (i in seq_len(n)) {
    
    porcentaje <- max(0, min(100, porcentajes[i]))
    frac <- porcentaje / 100
    offset <- (i - 1) * (1 + espacio)
    
    shapes_list[[length(shapes_list) + 1]] <- list(
      type = "rect",
      x0 = offset, x1 = offset + 1, y0 = 0, y1 = 1,
      fillcolor = "gray90",
      line = list(color = "black", width = 2)
    )
    
    shapes_list[[length(shapes_list) + 1]] <- list(
      type = "rect",
      x0 = offset + 0.5 - frac/2, x1 = offset + 0.5 + frac/2,
      y0 = 0.5 - frac/2, y1 = 0.5 + frac/2,
      fillcolor = "red",
      line = list(width = 0)
    )
    
    annotations_list[[length(annotations_list) + 1]] <- list(
      x = offset + 0.5, y = 0.5, text = paste0(round(porcentaje), "%"),
      showarrow = FALSE, font = list(size = 17, color = "black")
    )
    
    annotations_list[[length(annotations_list) + 1]] <- list(
      x = offset + 0.5, y = -0.2, text = etiquetas[i],
      showarrow = FALSE, font = list(size = 14, color = "black")
    )
  }
  
  plot_ly() %>%
    layout(
      title = titulo,
      xaxis = list(range = c(-0.3, (n - 1) * (1 + espacio) + 1.3),
                   showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE,
                   scaleanchor = "y"),
      yaxis = list(range = c(-0.4, 1.3),
                   showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
      shapes = shapes_list,
      annotations = annotations_list,
      margin = list(l = 20, r = 20, t = 50, b = 20)
    )
}

prom_uti_pallet = mean(catalogo_especificaciones_operaciones$utilizacion)

cuadrados_porcentaje(
  porcentajes = prom_uti_pallet * 100,
  etiquetas = "Total",
  titulo = list(
    text = "Utilización promedio del pallet",
    font = list(size = 17, color = "black"),
    x = 0.5
  )
)
```

To embed an image from an external file, you can use the Insert menu in the Visual Editor in RStudio and select Figure / Image. This will pop open a menu where you can browse to the image you want to insert as well as add alternative text or caption to it and adjust its size. In the visual editor you can also simply paste an image from your clipboard into your document and RStudio will place a copy of that image in your project folder.
